Ethnicity sensitive author disambiguation using semi-supervised learning

### 2016-04-29 paper/author_disambiguation.tex
kesw: final rewording
Signed-off-by: Mateusz Susik <mateuszsusik@gmail.com>

Using Recursive Feature Elimination \cite{guyon2002gene}, we next evaluate the
usefulness of all fifteen standard and seven additional ethnicity features for learning
the linkage function. The analysis consists in using the state-of-the-art algorithm
first using all twenty two features, to determine the least discriminative from feature
importances \cite{louppe2013understanding}, and then re-learn the state-of-the-art algorithm
using all but that one feature. That process is repeated recursively until
eventually only one feature remains. Results are presented in
Figure~\ref{fig:rfe} for one of the three folds with
the state-of-the-art, starting from the far right, \textit{Second given name} being the least important feature,
and ending on the left with all features eliminated but \textit{Chinese}. As
the figure illustrates, the most important features are ethnic-based features
(\textit{Chinese}, \textit{Other Asian}, \textit{Black}) along with
\textit{Co-authors}, \textit{Affiliation} and \textit{Full name}. Adding the remaining
other features only brings marginal improvements, with \textit{Journal},
\textit{Abstract}, \textit{Collaborations}, \textit{References}, \textit{Given
name initial} and \textit{Second given name} being almost insignificant.
Overall, these results highlight the added value of the proposed ethnicity
features.
Their duality in modeling both the similarity between author names
and their origins make them very strong predictors for author disambiguation.
The results also corroborate those from \cite{kang2009co} or \cite{ferreira2010effective}, who
found that the similarity between co-authors was a highly discriminative
feature.
If computational complexity is a concern, this analysis also
shows how decent performance can be achieved using only a very
small set of features, as also observed in
\cite{treeratpituk2009disambiguating} or \cite{levin2012citation}.

@@

In this work, we have revisited and validated the general author disambiguation
pipeline introduced in previous independent research work.
The generic approach is composed of three components, whose design and tuning are all critical
to good performance: (i) a blocking function for pre-clustering signatures
and reducing computational complexity, (ii) a linkage function for identifying
signatures with coreferring authors and (iii) the agglomerative clustering of
signatures. Making use of a distinctively large dataset of more than 1 million
crowdsourced annotations, we experimentally study all three components and
propose further improvements. With regards to blocking, we suggest to use
phonetization of author names to increase recall while maintaining low
computational complexity. For the linkage function, we introduce
ethnicity-sensitive features for the automatic tailoring of disambiguation to non-Western
author names whenever necessary. Finally, we explore semi-supervised cut-off
threshold strategies for agglomerative clustering. For all three components,
experiments show that our refinements all yield significantly better author
disambiguation accuracy.
In general, the results encourage further improvements and research. For blocking, one of the challenges is to manage signatures with inconsistent surnames or first given names (cases 4 and 5, as described in Section~\ref{methods:blocking}) while maintaining blocks to a tractable size. As phonetic algorithms are not yet perfect, another direction  for further work is the design of better phonetization functions, tailored for author disambiguation. For the linkage function, the good results of the proposed features pave the way for further research in ethnicity-sensitive author disambiguation. The automatic fitting of the pipeline to ethnic groups is better that a standard one that is less efficient for author names with many homonyms (\eg,Chinese authors). Collecting a bigger name-to-ethnicity dataset with better coverage of names is another future work to consider.
Moreover, the techniques presented in this work can be easily adapted to the broader
problem of named entity disambiguation and might improve accuracy of semantic search
algorithms.

@@

\verylongpage

>>>>

Using Recursive Feature Elimination \cite{guyon2002gene}, we next evaluate the
usefulness of all fifteen standard and seven additional ethnicity features for learning
the linkage function. The analysis consists in using the state-of-the-art algorithm
first using all twenty two features, to determine the least discriminative from feature
importances \cite{louppe2013understanding}, and then re-learn the state-of-the-art algorithm
using all but that one feature. That process is repeated recursively until
eventually only one feature remains. Results are presented in
Figure~\ref{fig:rfe} for one of the three folds with
the state-of-the-art, starting from the far right, \textit{Second given name} being the least important feature,
and ending on the left with all features eliminated but \textit{Chinese}. As
the figure illustrates, the most important features are ethnic-based features
(\textit{Chinese}, \textit{Other Asian}, \textit{Black}) along with
\textit{Co-authors}, \textit{Affiliation} and \textit{Full name}. Adding the remaining
other features only brings marginal improvements.
Overall, these results highlight the added value of the proposed ethnicity
features.
Their duality in modeling both the similarity between author names
and their origins make them very strong predictors for author disambiguation.
The results also corroborate those from \cite{kang2009co} or \cite{ferreira2010effective}, who
found that the similarity between co-authors was a highly discriminative
feature.
If computational complexity is a concern, this analysis also
shows how decent performance can be achieved using only a very
small set of features, as also observed in
\cite{treeratpituk2009disambiguating} or \cite{levin2012citation}.

@@

In this work, we have revisited and validated the general author disambiguation
pipeline introduced in previous independent research work.
The generic approach is composed of three components, whose design and tuning are all critical
to good performance: (i) a blocking function for pre-clustering signatures
and reducing computational complexity, (ii) a linkage function for identifying
signatures with coreferring authors and (iii) the agglomerative clustering of
signatures. Making use of a distinctively large dataset of more than 1 million
crowdsourced annotations, we experimentally study all three components and
propose further improvements. With regards to blocking, we suggest to use
phonetization of author names to increase recall while maintaining low
computational complexity. For the linkage function, we introduce
ethnicity-sensitive features for the automatic tailoring of disambiguation to non-Western
author names whenever necessary. Finally, we explore semi-supervised cut-off
threshold strategies for agglomerative clustering. For all three components,
experiments show that our refinements all yield significantly better author
disambiguation accuracy.
In general, the results encourage further improvements and research. For blocking, one of the
challenges is to manage signatures with inconsistent surnames or first given names (cases 4
and 5, as described in Section~\ref{methods:blocking}) while maintaining blocks to a tractable
size. As phonetic algorithms are not yet perfect, another direction  for further work is the
design of better phonetization functions, tailored for author disambiguation. For the linkage
function, the good results of the proposed features pave the way for further research in
ethnicity-sensitivity. The automatic fitting of the
pipeline to cultures and ethnic groups for which standard author disambiguation
is known to be less efficient (\eg, Chinese authors with many homonyms) indeed
constitutes a direction of research with great potential benefits for the
concerned scientific communities. Exploring other name-to-ethnicity datasets with deeper 
coverage of names is another future work worth considering.

Moreover, the techniques presented in this work can be easily adapted to the broader
problem of named entity disambiguation and thus might significantly improve accuracy of
semantic search algorithms.

@@
